---
title: "Google Trends Work"
output: html_notebook
---

This document will detail the data cleaning for the Google Trends project that is being led in large part by Alex Cooper based on his work at Child care aware of America.

This work was submitted and presented at the Symposium on Spatiotemporal Data Science in Washington DC in July of 2025. The work has been submitted and accepted for a paper presentation in July of 2025. The paper is currently under review in the journal "Papers in Applied Geography".


First we have to load the necessary packages and load in the data needed for the analysis. The data comes from pulls from pytrends. Please see the .readme in the github repsoitory for more information.

```{r}
## Load in the necessary packages
library(dplyr)
library(ggplot2)
library(tidyr)
library(sf)
library(maps)
library(tidycensus)
library(tidyverse)
library(readxl)
library(moments)

## Read in the functions file for this project
source("./google_trends_functions.R")
```



## United States

Let's first take a look at data for the entire United States. There are several files associated with this US data.

1. Summary file
2. Time dataset

These files are available at various geographic levels

1. Entire country
2. States
3. DMA

We are going to take a look at this data and the variability across the datasets across time and by geography

### National Level data

```{r}
## Load in the data
us_raw <- read_excel("./data/usa_temporal_df_20180603-20220910_raw_data_records.xlsx")
```

What are the descriptive statistics that we are going to calculate?
* Mean
* Standard Deviation
* Variance
* Min & Max
* Skewness
* Kurtosis

```{r}
## Some quick data cleaning

## convert strings to dates
# us_summary$event_time <- as.Date(us_summary$event_time)
us_raw$pull_date <- as.Date(us_raw$pull_date, format = "%m/%d/%Y")

## Remove the NA values
us_raw <- na.omit(us_raw) # Really only need this because there are dates that are missing and the dataframe won't build. I removed it from the other code chunks
```


```{r}
## Generate some descriptive statistics for the national level data

us_descriptive <- get_descriptive_stats_time(us_raw)

## add time column
us_descriptive <- add_date_column(us_descriptive)
```



### State Level Data

```{r}
## Load in the data
us_state_raw <- read_excel("./data/usa_states_df_20180603-20220910_raw_data_records.xlsx")
# us_state_summary <- read_excel("/Users/seanreid/Library/CloudStorage/Box-Box/google_trends/data/googletrendsdatasetsasofjanuary222024/usa_temporal_df_20180603-20220910_raw_data_records.xlsx")
```

```{r}
## Some quick data cleaning

## convert strings to dates
# us_state_summary$event_time <- as.Date(us_state_summary$event_time)
us_state_raw$pull_date <- as.Date(us_state_raw$pull_date, format = "%m/%d/%Y")

## Remove the NA values
# us_state_raw <- na.omit(us_state_raw)
```

```{r}
## Get descriptive stats for each of the states
us_state_descriptive <- get_descriptive_stats_time(us_state_raw)

## Add a geography column
us_state_descriptive <- add_geog_column(us_state_descriptive)
```




### DMA Level Data

```{r}
## Load in the data

us_raw_dma <- read_excel("./data/usa_dma_df_20180603-20220910_raw_data_records.xlsx")
```

```{r}
## some quick data cleaning

## Convert strings to dates
us_raw_dma$pull_date <- as.Date(us_raw_dma$pull_date, format = "%m/%d/%Y")

## Remove NA values
## Hmmm this isn't working and is zeroing out all the values
# us_raw_dma <- na.omit(us_raw_dma) # Commenting this out for now
```

```{r}
## Get descriptive stats for each of the DMA areas
us_dma_descriptive <- get_descriptive_stats_time(us_raw_dma)

## Add geog column
us_dma_descriptive <- add_geog_column(us_dma_descriptive)
```






## Oregon

```{r}
oregon_raw <- read_excel("./data/or_dma_20200214-20210214_raw_data_records.xlsx")
oregon_time <- read_excel("./data/or_time_20200214-20210214_raw_data_records.xlsx")
# oregon_summary <- read_excel("/Users/seanreid/Library/CloudStorage/Box-Box/google_trends/data/googletrendsdatasetsasofjanuary222024/or_dma_20200214-20210214_summary_stats.xlsx")

eugene_time <- read_excel("./data/eugene_time_20200214-20210214_raw_data_records.xlsx")
```

```{r}
## Do a little bit of data cleaning

oregon_time$pull_date <- as.Date(oregon_time$pull_date, format = "%m/%d/%Y")

## Remove the NA values
# oregon_time <- na.omit(oregon_time)

## Get the descriptive stats
oregon_time_descriptive <- get_descriptive_stats_time(oregon_time)

## Add the time column
oregon_time_descriptive <- add_date_column(oregon_time_descriptive)
```

```{r}
## Do the same thing for the raw level data. Remember, you don't have to do the na.omit command here

oregon_raw$pull_date <- as.Date(oregon_raw$pull_date, format = "%m/%d/%Y")

## Get the descriptive statistics
oregon_raw_descriptive <- get_descriptive_stats_time(oregon_raw)

## Add a geog column
oregon_raw_descriptive <- add_geog_column(oregon_raw_descriptive)
```

```{r}
## Clean up the Eugene data

## Format the date column correclty
eugene_time$pull_date <- as.Date(eugene_time$pull_date, format = "%m/%d/%Y")

## Omit missing days
# eugene_time <- na.omit(eugene_time)

## Get descriptives
eugene_time_descriptive <- get_descriptive_stats_time(eugene_time)

## Add a time column
eugene_time_descriptive <- add_date_column(eugene_time_descriptive)
```


## Indiana

```{r}
## Load in the data
indiana_time <- read_excel("./data/in_time_20180603-20220910_raw_data_records.xlsx")
indiana_dma <- read_excel("./data/in_dma_20180603-20220910_raw_data_records.xlsx")
```

```{r}
## Do a little bit of data cleaning

indiana_time$pull_date <- as.Date(indiana_time$pull_date, format = "%m/%d/%Y")

## Remove the NA values
# indiana_time <- na.omit(indiana_time)

## Get the descriptive stats
indiana_time_descriptive <- get_descriptive_stats_time(indiana_time)

## Add the time column
indiana_time_descriptive <- add_date_column(indiana_time_descriptive)

indiana_dma$pull_date <- as.Date(indiana_dma$pull_date, format = "%m/%d/%Y")

## Get the descriptive statistics
indiana_dma_descriptive <- get_descriptive_stats_time(indiana_dma)

## Add a geog column
indiana_dma_descriptive <- add_geog_column(indiana_dma_descriptive)
```



## Minnesota

```{r}
## Load in the data
mn_dma <- read_excel("./data/mn_dma_20200214-20210214_raw_data_records.xlsx")
mn_time <- read_excel("./data/mn_time_20200214-20210214_raw_data_records.xlsx")
```


```{r}
## Do a little bit of data cleaning

mn_time$pull_date <- as.Date(mn_time$pull_date, format = "%m/%d/%Y")

## Remove the NA values
# mn_time <- na.omit(mn_time)

## Get the descriptive stats
mn_time_descriptive <- get_descriptive_stats_time(mn_time)

## Add the time column
mn_time_descriptive <- add_date_column(mn_time_descriptive)



mn_dma$pull_date <- as.Date(mn_dma$pull_date, format = "%m/%d/%Y")

## Get the descriptive statistics
mn_dma_descriptive <- get_descriptive_stats_time(mn_dma)

## Add a geog column
mn_dma_descriptive <- add_geog_column(mn_dma_descriptive)
```


## Ohio

```{r}
## Load in the data
oh_dma <- read_excel("./data/oh_dma_20180603-20220910_raw_data_records.xlsx")
oh_time <- read_excel("./data/oh_time_20180603-20220910_raw_data_records.xlsx")
```

```{r}
## Do a little bit of data cleaning

oh_time$pull_date <- as.Date(oh_time$pull_date, format = "%m/%d/%Y")

## Remove the NA values
# oh_time <- na.omit(oh_time)

## Get the descriptive stats
oh_time_descriptive <- get_descriptive_stats_time(oh_time)

## Add the time column
oh_time_descriptive <- add_date_column(oh_time_descriptive)



oh_dma$pull_date <- as.Date(oh_dma$pull_date, format = "%m/%d/%Y")

## Get the descriptive statistics
oh_dma_descriptive <- get_descriptive_stats_time(oh_dma)

## Add a geog column
oh_dma_descriptive <- add_geog_column(oh_dma_descriptive)
```


## Kentucky

```{r}
## Load in the data
ky_time <- read_excel("./data/ky_time_20180603-20220910_raw_data_records.xlsx")
```

```{r}
ky_time$pull_date <- as.Date(ky_time$pull_date, format = "%m/%d/%Y")

## Get the descriptive stats
ky_time_descriptive <- get_descriptive_stats_time(ky_time)

## Add the time column
ky_time_descriptive <- add_date_column(ky_time_descriptive)

```



## Texas

```{r}
## Load in the data
tx_dma <- read_excel("./data/tx_dma_20210321-20210421_raw_data_records.xlsx")
tx_time <- read_excel("./data/tx_time_20210321-20210421_raw_data_records.xlsx")
```

```{r}
## Do a little bit of data cleaning

tx_time$pull_date <- as.Date(tx_time$pull_date, format = "%m/%d/%Y")

## Remove the NA values
# tx_time <- na.omit(tx_time)

## Get the descriptive stats
tx_time_descriptive <- get_descriptive_stats_time(tx_time)

## Add the time column
tx_time_descriptive <- add_date_column(tx_time_descriptive)




tx_dma$pull_date <- as.Date(tx_dma$pull_date, format = "%m/%d/%Y")

## Get the descriptive statistics
tx_dma_descriptive <- get_descriptive_stats_time(tx_dma)

## Add a geog column
tx_dma_descriptive <- add_geog_column(tx_dma_descriptive)
```



## Plotting the data

To effectively plot how the data is different over time based on geography I am going to combine all the time data into the same dataframe so it's easier to work with

```{r}
## Reset the index values for the dataframes
row.names(us_descriptive) <- NULL
row.names(oregon_time_descriptive) <- NULL
row.names(eugene_time_descriptive) <- NULL
row.names(mn_time_descriptive) <- NULL
row.names(oh_time_descriptive) <- NULL
row.names(indiana_time_descriptive) <- NULL
row.names(tx_time_descriptive) <- NULL
row.names(ky_time_descriptive) <- NULL

## Add a column to each of the dataframes so I know the geography they are coming from
us_descriptive$source <- "US"
oregon_time_descriptive$source <- "Oregon"
eugene_time_descriptive$source <- "Eugene"
mn_time_descriptive$source <- "Minnesota"
oh_time_descriptive$source <- "Ohio"
indiana_time_descriptive$source <- "Indiana"
tx_time_descriptive$source <- "Texas"
ky_time_descriptive$source <- "Kentucky"

## Bind the 3 dataframes together
time_data <- rbind(us_descriptive, oregon_time_descriptive, eugene_time_descriptive, mn_time_descriptive, oh_time_descriptive, indiana_time_descriptive, tx_time_descriptive, ky_time_descriptive)

## Pinch the data so that it's on the same time scale
# time_data <- filter(time_data, time %in% oh_time_descriptive$time)
```


At this point the data cleaning is complete and we are ready to run the analysis. Move to the google_trends_analysis.Rmd file to complete the next steps!
